Cooperative Kernels: GPU Multitasking for Blocking Algorithms

### 2017-07-06 write-ups/FSE/paper.tex
This is such a good paper.

\myparagraph{Execution Model}
OpenCL~\cite[p.\ 31]{opencl2Spec} and CUDA~\cite{cuda-75} specifically makes no guarantees about fair scheduling between
workgroups executing the same kernel. 
%stating~\cite[p.\ 31]{opencl2Spec}: \emph{``A
%  conforming implementation may choose to serialize the workgroups
%so
%  a correct algorithm cannot assume that workgroups will execute in
%  parallel.
%\dots There is no safe and portable way to synchronize across
%  the independent execution of workgroups since once in the work-pool,
%  they can execute in any order.''}  CUDA similarly provides no guarantees~\cite{cuda-75}.
%
HSA provides limited, one-way guarantees,
stating~\cite[p. 46]{HSAprogramming11}: \emph{Work-group A can wait
  for values written by work-group B without deadlock provided ... (if) A
  comes after B in work-group flattened ID order}. This is not sufficient to support blocking algorithms that use
mutexes and inter-workgroup barriers, both of which require \emph{symmetric} communication between
threads.

@@

If a task was obtained, then the workgroup processes it
(line~\ref{line:wksteal:processtask}), which may lead to new tasks
being created and pushed to the workgroup's queue. The kernel presents
two opportunities for spin-waiting: spinning to obtain a mutex, and
spinning in the main kernel loop to obtain a task. \TSAdded{Without fair
scheduling, threads waiting for the mutex might spin indefinitely,
causing the application to hang.}

@@

The application iterates as long as the current frontier contains
nodes to process (line~\ref{line:graph:iterate}). At each frontier,
the nodes to be processed are evenly distributed between
threads through \emph{stride} based processing.
%
In this case, the stride is the total number of threads, obtained via
$\getglobalsize$.  
%Each thread processes nodes with indices
%\code{$\keyword{tid}$+$\keyword{stride}$*0},
%\code{$\keyword{tid}$+$\keyword{stride}$*1},
%\code{$\keyword{tid}$+$\keyword{stride}$*2}, etc., where $\keyword{tid}$ is a thread's global id.
A thread calls $\keyword{process\_node}$ to process a node given the current level, with nodes to be processed next frontier being pushed to $\keyword{out\_nodes}$. After processing the frontier, the threads swap their
node array pointers (line~\ref{line:graph:swap}).

@@

The global barrier used in this application is not provided as a GPU
primitive, though previous works have shown that such a global barrier
can be implemented~\cite{XF10,DBLP:conf/oopsla/SorensenDBGR16}, based
on CPU barrier designs~\cite[ch. 17]{HS08}.  These barriers employ
spinning to ensure threads wait at the barrier until all threads have
arrived, which requires fair scheduling between workgroups for the
barrier to operate correctly. \TSAdded{Without fair scheduling, the
  barrier threads may wait indefinitely at the barrier, causing the
  application to hang.}

>>>>

\myparagraph{Execution Model}
OpenCL~\cite[p.\ 31]{opencl2Spec} and CUDA~\cite{cuda-75} specifically make no guarantees about fair scheduling between
workgroups executing the same kernel. 
%stating~\cite[p.\ 31]{opencl2Spec}: \emph{``A
%  conforming implementation may choose to serialize the workgroups
%so
%  a correct algorithm cannot assume that workgroups will execute in
%  parallel.
%\dots There is no safe and portable way to synchronize across
%  the independent execution of workgroups since once in the work-pool,
%  they can execute in any order.''}  CUDA similarly provides no guarantees~\cite{cuda-75}.
%
HSA provides limited, one-way guarantees,
stating~\cite[p. 46]{HSAprogramming11}: \emph{``Work-group A can wait
  for values written by work-group B without deadlock provided ... (if) A
  comes after B in work-group flattened ID order''}. This is not sufficient to support blocking algorithms that use
mutexes and inter-workgroup barriers, both of which require \emph{symmetric} communication between
threads.

@@

If a task is obtained, then the workgroup processes it
(line~\ref{line:wksteal:processtask}), which may lead to new tasks
being created and pushed to the workgroup's queue. The kernel presents
two opportunities for spin-waiting: spinning to obtain a mutex, and
spinning in the main kernel loop to obtain a task. \TSAdded{Without fair
scheduling, threads waiting for the mutex might spin indefinitely,
causing the application to hang.}

@@

The application iterates as long as the current frontier contains
nodes to process (line~\ref{line:graph:iterate}). At each frontier,
the nodes to be processed are evenly distributed between
threads through \emph{stride} based processing.
%
In this case, the stride is the total number of threads, obtained via
$\getglobalsize$.  
%Each thread processes nodes with indices
%\code{$\keyword{tid}$+$\keyword{stride}$*0},
%\code{$\keyword{tid}$+$\keyword{stride}$*1},
%\code{$\keyword{tid}$+$\keyword{stride}$*2}, etc., where $\keyword{tid}$ is a thread's global id.
A thread calls $\keyword{process\_node}$ to process a node given the current level, with nodes to be processed during the next frontier being pushed to $\keyword{out\_nodes}$. After processing the frontier, the threads swap their
node array pointers (line~\ref{line:graph:swap}).

@@

The global barrier used in this application is not provided as a GPU
primitive, though previous works have shown that such a global barrier
can be implemented~\cite{XF10,DBLP:conf/oopsla/SorensenDBGR16}, based
on CPU barrier designs~\cite[ch. 17]{HS08}.  These barriers employ
spinning to ensure each thread waits at the barrier until all threads have
arrived, thus fair scheduling between workgroups is required for the
barrier to operate correctly. \TSAdded{Without fair scheduling, the
  barrier threads may wait indefinitely at the barrier, causing the
  application to hang.}
